### 2021-02-18 paper/linkagecluster_bk_paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Linkage Cluster Paper Draft}

    Brian Kundinger

    \end{center}

\hypertarget{related-work}{%
\subsection{Related Work}\label{related-work}}

Sadinle (2017) proposed a Bayesian method for bipartite record linkage
that provides a foundation for much of our proposed method. In order to
ensure one-to-one matching, his Gibbs sampler conditions on the linkage
status of all other records when making each linkage assignment. While
this elegantly ensures transitivity, the probability distribution of the
linkage vector \(\mathbf{Z}\) is somewhat complicated, and does not
easily allow for inclusion of additional structure among records. Our
method however, motivated by recent decouple/recouple methods, initially
relaxes the one-to-one matching assumption in order to incorporate
linkage information differing by subgroups.

There has been some work done in the literature on conducting record
linkage in situations where feild accuracy differs by subgroups of the
data, but these are hampered by technical issues that make them
difficult to use in practice. Larson (2005) proposed Bayesian
heirarchical record linkage model that explicitly modelled the
differences in parameters across blocks of the data. In relying on
blocking however, he assumes that the blocking variable is error-free,
when in practice such a procedure could result in many missed linkages.
In contrast, our proposed method induces high probabilities of matching
within a linkage cluster while still allowing the possibility to match
across clusters (when the clustering field is recorded with error. Hu et
al (2015) proposed a method that used a logistic regression submodel
within a broader record linkage procedure to account for ``expected
changes'' in the record of the underlying entity based on subgroup
membership. This method however specified for use in temporal settings
when the true attributes of the record change with time, and it is
unclear how such a method would be implemented when records are
collected syncronously with error, or when time information is
unavailable. Additionally, the method is semi-supervised, requiring
training data to implement the logistic regression, and is that not well
suited for record linkage in practice. In contrast, the proposed method
is fully unsupervised and is shown to not require extensive tuning of
prior distributions, allowing for ease of implementation in a variety of
settings.

\hypertarget{notation-and-assumptions}{%
\subsection{Notation and Assumptions}\label{notation-and-assumptions}}

Our notation and assumptions closely follow that of Sadinle (2017).
Denote two files as \(\mathbf{X}_A\) and \(\mathbf{X}_B\), with \(n_A\)
and \(n_B\) records respectively, and with records indexed as
\(i \in \{1, \ldots, n_A\}\) in \(\mathbf{X}_A\) and
\(j \in \{1, \ldots, n_B\}\) in \(\mathbf{X}_B\). Without loss of
generality, label the files such that \(n_A \geq n_B\). We also assume
there are no duplicates within files, only across. For each record pair
under consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each takes
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating level agreement between the two records on a specified field.

Following the Fellegi Sunter framework, we define
\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
observing agreement level \(l\) in field \(f\) for records \(i\) and
\(j\) given that the records are a match, and similarly define
\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We
denote \(\lambda\) to be the (marginal) probability that some record
\(j \in \mathbf{X}_B\) has a match in \(\mathbf{X}_A\).

Lastly, we adopt from Sadinle 2017 the matching labelling
\(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\), defined as \[Z_j=\begin{cases} 
    i,  & \text{if records } i\in \mathbf{X}_A \text{ and } j\in \mathbf{X}_B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in \mathbf{X}_B \text{ does not have a match in file} \mathbf{X}_B; \\
\end{cases}\]

Wherever possible, we reserve superscripts for denoting field, level,
and linkage cluster, and reserve subscripts for record indices. For
example, \(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the
probability distribution governing field \(f\) for matching records, and
\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
is product of the relevant of the appropriate \(\mathbf{m}\) parameters
for record pair \((i,j)\). We hope that these conventions avoid
overloaded notation in the likelihood and subsequent derivations.

\hypertarget{model-specification}{%
\section{Model Specification}\label{model-specification}}

Additionally, our model differs from that of Sadinle 2017 through its
explicit dependence on a beta random variable \(\lambda\) that models
the rate of matching across records. Sadinle marginalizes over such a
random variable in his derivations of the ``beta prior for bipartite
matching,'' but here we provide derivations without marginalizing in
order to specify differing rates of matching for different linkage
clusters.

\hypertarget{prior-distributions-and-likelihood}{%
\subsection{Prior Distributions and
Likelihood}\label{prior-distributions-and-likelihood}}

For fields \(f \in \{1, \ldots, F\}\) and levels
\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
distributions. Note that likelihood is related to that of Sadinle
(2017), but includes dependence on the \(\lambda\) random variable. It
also makes it explict that units of observation under this model are the
\(n_B\) records in \(\mathbf{X}_B\), a crucial distinction when compared
to the likelihood over the \(n_A \times n_B\) record pairs in the
original Fellegi Sunter model.

\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \left[\lambda^{\mathbf{1}_{z_j < n_A + 1}} (1-\lambda)^{\mathbf{1}_{z_j = n_A + 1}}\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\right]\]

\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
\[Z_j | \lambda =
\begin{cases} 
    \frac{1}{n_A}\lambda  & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in \mathbf{X}_A\), and non-matching probability governed by
\(\lambda\). Observe two reasonable choices for \(p(Z, \lambda)\):

If \(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\), then the prior for
\(Z\), integrated over \(\lambda\), becomes

\[=\begin{cases} 
    1  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\] Thus this is a uniform prior on the labellings of \(Z\). A
second useful prior would be \(\lambda \sim \text{Beta}(1, 1)\), which
yields \[=\begin{cases} 
    1  & z_j \leq n_A; \\
    n_A &  z_j  = n_A + 1 \\
\end{cases}\] Here, a match and a non-match are equaly likely, and among
the matches, all records are likely to be a match. There is little
difference between the the two priors in practice, as the data should
overpower the prior.

\hypertarget{posterior-sampling}{%
\subsection{Posterior Sampling}\label{posterior-sampling}}

The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
standard multinomial-dirichlet mechanics. Thus we have
\[\mathbf{m^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
\[\mathbf{u^{f}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
and
\(\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).

Since \(\lambda\) provides the probability that a record in
\(\mathbf{X}_B\) has match in \(\mathbf{X}_A\), it has a full
conditional distribution

\[\lambda|\mathbf{Z^{(s)}} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]
where \(D = \sum_1^{n_B} (Z_j < n_A + 1)\) is the number of records
identified as matches in the previous Gibbs iteration.

Lastly, we sample each \(Z_j\) according to these posterior
distributions. To do so, note that as a function of \(Z_j\), the
likelihood is a discrete distibution with probabilities proportional to

\[p(\Gamma_{.j}|Z_{j}) \propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}} + \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\prod_{i=1}^{n_A}\mathbf{u}_{ij}\]
\[\propto \mathbf{1}_{z_j < n_A + 1}\lambda\prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}}+ \mathbf{1}_{z_j = n_A + 1}(1-\lambda)\]

\[\propto \lambda \left[\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i}\right] + (1 - \lambda)\mathbf{1}_{z_j = n_A + 1}\]
where
\(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
The interested reader should note that these are precisely the
likelihood ratios used in the Fellegi-Sunter model to classify matches
and non-matches, and we therefore refer to \(w_{ij}\) as the
\emph{Fellegi Sunter weights}.

The full conditional for \(Z_j\) is thus given by
\[p(Z_j^{(s+1)}|\mathbf{m}, \mathbf{u}, \Gamma, \lambda, \mathbf{Z^{(s)}}) \propto p(\Gamma_{.j}|\mathbf{m}, \mathbf{u}, Z_j, \lambda)\overbrace{p(\mathbf{m}, \mathbf{u}|\mathbf{Z^{(s)}})}^{\text{proportionality}}p(Z_j | \lambda) \overbrace{p(\lambda|\mathbf{Z^{(s)}})}^{\text{proportionality}}\]
\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
\begin{cases} 
    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

Here, one should note that if we choose to integrate over the posterior
of \(\lambda\) and rearrange terms, we produce the following sampling
mechanism:

\[Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
\end{cases}\]

The relationship between this sampling mechanism and that of Sadinle
2017 warrants discussion.

\hypertarget{bayes-estimate}{%
\subsection{Bayes Estimate}\label{bayes-estimate}}

Sadinle 2017

\hypertarget{decouplerecouple}{%
\subsection{Decouple/Recouple}\label{decouplerecouple}}

One difficulty in record linkage is conducting linkage while taking
appropriate dependencies into account. In the context of this paper in
which we assume that each file is recorded without duplicates, this
means our linkage procedure should respect one-to-one matching. The
posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
this issue; since we only sample one potential match for record \(j\) in
each Gibbs iteration, it is impossible to match \(j\) to two records at
once. This in itself is an improvement over the classic Fellegi-Sunter
method that labels record pairs as matches or non matches independently
of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
\(j' \neq j\), which cleary violates the assumption that
\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
this issue by adapting the set of candidate matches for every \(Z_j\),
but we forgo this rigidity and instead ``decouple'' the linkage vector
\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
independently. In our formulation, such illegal matches occasionally
occur within individual Gibbs iterations due to simple randomness, but
they do not occur consistently enough between the same records to
measurably influence the eventual Bayes estimate of the linkage
structure.

However, if we want to ensure the theoretic purity of each MCMC sample
of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
resolving these illegal matchings. Explictly, define clusters
\(C(i) := \{j | Z_j = i\}\), and define all clusters such that
\(|C(i)| > 1\) as \emph{conflict clusters}. Within each cluster,
resample the candidate matches according to their Fellegi-Sunter
weights. We show this step to be computationally inexpensive and to be a
reasonable replication of more organic one-to-one matching. This
procedure would be most useful in cases where there are many ``near
matches'' across files, but in most cases, we find this this step to be
unnecessary.

\hypertarget{linkage-cluster-model}{%
\subsection{Linkage cluster model}\label{linkage-cluster-model}}

The primary contribution of this paper is the establishment of linkage
clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
discernment parameters \(\mathbf{u}^f\), and also the linkage
probability parameter \(\lambda\) to vary according to certain
attributes in the record pairs. We shall see that this set up upweights
linkage within clusters and downweights linkage across clusters, thus
using attribute information in way not yet explored in the literture.

Assign each record in both files to a linkage cluster
\(S_k \in \{S_1, \ldots, S_K\}\) based on some attribute in the data. We
see that the comparison vector

\end{document}