### 2021-04-19 paper/FAQ_STA561_Project.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Parlr: Frequently Asked Questions}

    Brian Kundinger

    \end{center}

\hypertarget{what-are-parlrs-competitors}{%
\subsubsection{What are parlr's
competitors?}\label{what-are-parlrs-competitors}}

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their model was to transform the sets of records,
which constitute text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates a
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f\) providing their similarity score on the
\(f^{th}\) field. For ease of modeling and computation, we restrict
these similarity scores to be discrete, ordinal variables, and the
construction of these is left to the modeller. It is common to use use
binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
provide an option for partial matching. For text data, we calculate
similarity based on Levenstein distance or some other text similarity
score, and bin these scores to integers for use in the model.

Two modern adaptations of the Fellegi Sunter model are important for
understanding \texttt{parlr} and its contribution. In 2019, Enamorado et
al proposed \texttt{fastlink}, a method that closely followed the
modelling assumptions of Fellegi and Sunter, but used innovating hashing
techniques to greatly enhance its speed and scale. Importantly, they
maintained Fellegi and Sunter's independent matching assumption, in
which the matching status of any \((i,j)\) pair is made independently of
the matching of any other \((i', j')\). This leads to many matchings
that violate ``one-to-one'' considerations that are often implicit in
the data, which need to be resolved in a post-processing step. In
contrast, Sadinle (2017) proposed ``Beta Record Linkage for Bipartite
Matching'' and the accompanying \texttt{BRL} package which strictly
enforces one-to-one matching. Specifically, in each iteration of his
Gibbs sampler, he considers each record \(j\in B\), removes from
consideration the records \(i\in A\) that have already been matched,
then samples a potential link. However, accounting for these
dependencies throughout the linkage process is computionally burdensome,
leaving \texttt{BRL} only suitable for small to moderate linkage
problems.

\hypertarget{how-is-parlr-different}{%
\subsubsection{How is parlr different?}\label{how-is-parlr-different}}

While \texttt{fastlink} makes \(n_A \times n_B\) many indepedent binary
linkage decisions and Sadinle makes \(n_B\) many \emph{dependent}
linkage decisions among \(n_A\) many options, \texttt{parlr} strikes
middle ground by making \(n_B\) many \emph{independent} linkage
decisions among \(n_A\) many options. That is, in each iteration of the
Gibbs sampler, we make one linkage decision per \(j \in B\), but do not
remove previously matched records from consideration. While this does
mean that a particular iteration of the Gibbs sampler may contain some
matchings that violate one-to-one requirements, these violations happen
infreqently and randomly across all record pairing, so that they do not
seriously harm the final Bayes estimate of the linkage structure we make
from all the Gibbs samples. In cases where our Bayes estimate does
contain matches that violate one-to-one, we note that such cases far
less frequently than under \texttt{fastlink}, and that these cases are
easily resolved by simply taking the match pair with the highest
posterior match probability.

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(P\)
denote the total number of unique agreement patterns; note that \(P\) is
bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
scale with \(n_A\) or \(n_B\). We index these agreement patterns by
\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
conduct calculations over these \(P\) agreement patterns rather than the
\(n_A \times n_B\) record pairs. While the original Fellegi Sunter
framework has computational complexity \(O(n_A \times n_B)\), these
modifications completely remove dependence on \(n_A\) throughout the
Gibbs sampler, reducing the complexity to \(O(n_B)\)

\hypertarget{how-accurate-is-it}{%
\subsubsection{How accurate is it?}\label{how-accurate-is-it}}

In his 2017 paper, Sadinle demonstrated the strength of his method by
running \texttt{BRL} on simulated pairs of files with differing levels
of errors and overlap. Comparing our method against \texttt{BRL} on
these same simulated datasets, we find that our method only has weakend
performance in the extreme scenario of very high errors and very high
overlap across files. See plots in technical appendix.

\hypertarget{how-fast-is-it}{%
\subsubsection{How fast is it?}\label{how-fast-is-it}}

We simulate \(\Gamma\) matrices of comparison vectors from multinomial
distributions meant to emulate the behavior of similarity scores across
first name, last name, and day, month, and and year of birth. We
simulate these data for different values of \(n_A\) and \(n_B\), and
compare the runtime of \texttt{parlr} against \texttt{BRL}. We see that
at low data size, \texttt{BRL} outperforms, but that \texttt{parlr} is
significantly faster at handling larger data. See plots in technical
appendix.

\hypertarget{how-fast-can-it-be}{%
\subsubsection{How fast can it be?}\label{how-fast-can-it-be}}

We note here that \texttt{BRL} is coded in C, which makes for unfair
comparison against \texttt{parlr}, currently only built in R.
Additionally, although \texttt{parlr} is amenable to parallezation, our
simulations thus far have only been run on a single core. Running
\texttt{parlr} in C++ with paralellization for the hashing step and
sampling the matching status of the record pairs should lead to even
more drastic results.

\hypertarget{what-about-deduplication-of-a-single-file}{%
\subsubsection{What about deduplication of a single
file}\label{what-about-deduplication-of-a-single-file}}

At the moment, \texttt{parlr} (and its \texttt{BRL} competitor) is only
able to handle bipartite matching across two files, meaning there is
duplication across, but not within, files. Though it seems similar to
linkage across files, deduplication within one file is suprisingly much
more difficult to model probabalistically. In particular, \texttt{parlr}
assumes that the linkage decision for record \(j \in B\) is independent
from all other \(j' \in B\), allowing us to propose a prior
\(\lambda \sim \text{Beta}(\alpha, \beta)\) for the probability of \(j\)
having some match in \(A\), and update that prior through standard
beta-binomial posterior updates. However, with only one file of \(n\)
records, we do not even have \(n\) independent linkage decisions to work
with! For example, if the sampler decides that record \(1\) matches with
record \(5\), there is no way to determine the linkage of record \(5\)
independently. Constructing fast, scalable methods for deduplication is
one avenue of future work.

\hypertarget{anything-else}{%
\subsubsection{Anything else?}\label{anything-else}}

Kundinger is particularly interested in linkage scenarios when the
reliability of the data (the level of error in the fields), and the
rates of matching, differ by subgroups in the data. The simplest case of
this occurs when matching historical records for women; due to marriage
conventions, the last name is a much less reliable identifier for women
than for men. In an upcoming extention, we propose \emph{linkage
clusters} that account these difficulties, while still maintaining the
computational advances from \texttt{parlr}.

\end{document}