### 2021-06-06 paper/parlr_paper_1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}

    Brian Kundinger, Rebecca Steorts, and Jerome Reiter

    \end{center}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Entity resolution (or record linkage) is the task of identifying
duplicate records across multiple data sources. This task is
straightforward when provided unique identifiers or highly reliable
information, but becomes difficult when data sources are plagued with
error. Additionally, many entity resolution tasks are large in nature,
and many existing methods are too computational intensive to be used in
practice. In this paper, we propose a Bayesian model for record linkage
that uses hashing techniques and efficient storage mechanisms, and is
amenable to parallelization, making it scalable to large linkage tasks.

The proposed method is inspired by the \emph{Beta Record Linkage}
(\texttt{BRL}) method proposed in Mauricio Sadinle's 2017 ``Bayesian
Estimation of Bipartite Matching for Record Linkage''. To recognize this
lineage, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable''). In what follows, Section 2 describes previous
work in record linkage that lays the foundation for our method, Section
3 provides our model and the Gibbs sampler used for posterior inference,
and Section 4 describes strategies used for more efficient and scalable
implementation. Then, Section 5 demonstrates the accuracy and speed of
out method through two simulation studies, and Section 6 reveals
interesting properties of the method through a case study of El
Salvadoran casuality records. Lastly, our discussion highlights open
questions in the record linkage liturature and motivates future work.

\hypertarget{related-work}{%
\subsection{Related Work}\label{related-work}}

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their approach was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates an
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
score on the \(f^{th}\) field. For ease of modeling and computation, we
restrict these similarity scores to be discrete, ordinal variables, and
the construction of these is left to the modeler. We adopt the
convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
is common to use use binary 1-2 variables to indicate exact matching,
and 1-2-3 variables to provide an option for partial matching. For text
data, we calculate similarity based on Levenstein distance or some other
text similarity score, and bin these scores to integers for use in the
model. Jaro (1989) provided guidelines for appropriate binning for the
comparison vectors, but details of implementation are generally context
specific.

The likelihood in the original Fellegi Sunter model is a mixture model
through which each record pair is independently classified as a match or
nonmatch. This independent classification often leads to sets of matches
that break expectations of transitivity, which is undesirable. Although
Jaro devised an procedure to reduce a set of conflicting matches to the
mostly likely set of one-to-one matchings, the abundance of false
matches can bias the estimation of other parameters of the model.

One key strength of Sadinle's \texttt{BRL} is that it is garunteed to
produce a set of bipartite matchings that respect one-to-one
assumptions--without the need for any post-hoc procedures. Specifically,
in each iteration of his Gibbs sampler, he considers each record
\(j\in B\), removes from consideration the records \(i\in A\) that have
already been matched, and then samples a potential link. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

Outside of the Fellegi-Sunter framework, some researchers have developed
methods that model the data directly, not a derived set of comparison
vectors. As an early example, Steorts et al 2016 produced
\texttt{blink}, which had the advantage of a computational complexity
that grew linearly with the data (as opposed to quadratically within
Fellegi Sunter), but was plagued with slow mixing times and was unable
to incorporate text data. Later, Steorts et al 2020 improved on this
method with \texttt{dblink}, which was able to simulate text data within
its Gibbs sampler by drawing from an empirical distribution, and
remedied its slow mixing through a probabilistic blocking approach.
However, this method remains computationally intensive and requires
distributed computing to manage even moderate linkage problems in
acceptable time. Because these methods fall outside of the
Fellegi-Sunter framework, we do not consider them further in this paper.

\hypertarget{notation-and-assumptions}{%
\subsection{Notation and Assumptions}\label{notation-and-assumptions}}

Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
generality, label the files such that \(n_A \geq n_B\). We assume there
are no duplicates within files, only across. For each record pair under
consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating the level agreement between the two records on a specified
field.

To indicate matching status, we adopt the \emph{linkage structure
parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
defined as \[Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}\] This provides more memory efficient storage for the
linkage information than a \(n_A \times n_B\) sparse matrix of
indicators.

Following the Fellegi Sunter framework, we define
\(m^{fl}:= P(\gamma_{ij}^f = l |Z_j = i)\) to be the probability of
observing agreement level \(l\) in field \(f\) for records \(i\) and
\(j\) given that the records are a match, and similarly define
\(u^{fl}:= P(\gamma_{ij}^f = l |Z_j \neq i)\), for non-matches. We also
adopt Fellegi and Sunter's conditionally independent fields assumption
that the level of agreement on one field is independent of the level of
agreement on another. Though this assumption is often not reasonable
(for example, first name and gender are two clearly dependent fields),
but it is common within the record linkage literature and generally
leads to models that perform well in practice; see discussion for
further remarks. Lastly, we define \(\lambda\) to be the (marginal)
probability that some record \(j \in B\) has a match in \(A\).

Wherever possible, we reserve superscripts for denoting field and level,
while reserving subscripts for record indices. For example,
\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
distribution governing field \(f\) for matching records, and
\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
is product of the relevant of the appropriate \(\mathbf{m}\) parameters
for record pair \((i,j)\). We hope that these conventions avoid
overloaded notation in the likelihood and subsequent derivations.

\hypertarget{model-specification}{%
\subsection{Model Specification}\label{model-specification}}

For fields \(f \in \{1, \ldots, F\}\) and levels
\(l\in \{1, \ldots, L_f\}\) we adopt the following likelihood and prior
distributions.

\[P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) =\prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]

\[\mathbf{m^{f}} \sim \text{Dirichlet}(\alpha^{f1}, \ldots, \alpha^{fL_f})\]
\[\mathbf{u^{f}} \sim \text{Dirichlet}(\beta^{f1}, \ldots, \beta^{fL_f})\]
\[Z_j | \lambda =
\begin{cases} 
    \frac{1}{n_A}\lambda  & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in A\), and non-matching probability governed by \(\lambda\).
Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
belief that nonmatches and matches are equally likely, and a
\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)\) prior
corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).

Here the reader should the relationship between our proposed model and
that of Sadinle 2017. In his model, Sadinle constructs a prior
distribution on the entire \(\mathbf{Z}\) vector, which induces a Gibbs
sampler that strictly enforces one-to-one matching. In particular, this
sampler removes previously matches records from the set of candidate
records when sampling \(Z_j\), creating a dependency that makes the
sampler \emph{inherently serial}. We however use independent priors for
each \(Z_j\), creating a sampler that is \emph{perfectly
parallelizable}, allowing for significant computational gains. More
importantly, since only the agreement pattern of \(Z_j\) is used for
calculations within the Gibbs sampler, and not the particular record
label, we can conduct this sampling only at the level of the unique
agreement patterns, offering even more computational savings. In doing
so however, we thereby weaken the one-to-one requirement from
\texttt{BRL}; our sampler does ensure that each record in \(B\) can be
matched to at most one record in \(A\), but allows for the possiblity
that multiple records in \(B\) match to the same record in \(A\). We
explore the ramifications of this distinction with the El Salvador case
study.

\hypertarget{posterior-sampling}{%
\subsection{Posterior Sampling}\label{posterior-sampling}}

We work with the following factorization of the joint distribution:

\[p(\Gamma, \mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) = p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]

This factorization leads to following Gibbs Sampler:

\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
standard multinomial-dirichlet mechanics. Thus we have

\[\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
\[\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
and
\(\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
determined only by the number of duplicates
\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
\(\mathbf{Z}\). Thus we have

\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
\[\propto \lambda^D (1-\lambda)^{n_B - D} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
\[ \propto \lambda^{D + \alpha_{\lambda} - 1} (1-\lambda)^{n_B - D + \beta_{\lambda} -1}\]
\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_B - D + \beta_{\lambda})\]

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
is a discrete distribution with probabilities proportional to

\begin{align*}
p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
&=
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\\
\end{align*}

where
\(w_{ij} = \frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
The interested reader should note that these are precisely the
likelihood ratios used in the Fellegi-Sunter model to classify matches
and non-matches, and we therefore refer to \(w_{ij}\) as the
\emph{Fellegi Sunter weights}.

With the likelihood in this form, we can derive the full conditional
\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]

\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
\begin{cases} 
    \frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
     1-\lambda &  z_j  = n_A + 1 \\
\end{cases}\]

In order to make fair comparisons against the Sadinle 2017 model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditional:

\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{ij}  & i \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
\end{cases}\]

\hypertarget{bayes-estimate}{%
\subsection{Bayes Estimate}\label{bayes-estimate}}

We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
parameter \(\mathbf{Z}\) by assigning different positive losses to
different types of errors, and minimizing posterior expected loss. We
adopt the loss function proposed in Sadinle 2017, in which
\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
option to leave the matching undetermined by the model. Specifically, we
minimize the quantity
\(L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)\)
where

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
    0  & \text{if } Z_j = \hat{Z_j}; \\
    \theta_R,  & \text{if } \hat{Z_j} = R; \\
    \theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
    \theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
    \theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model. This loss function leads to closed form decision rules for
minimizing posterior expected loss. In this paper, we adopt losses
\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\),
inducing the intuitive decision rule

\[\hat{Z}_j =\begin{cases} 
    i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
    0,  & \text{otherwise} ; \\
\end{cases}\]

For a more indepth explanation of this function and the induced Bayes
estimate, see Sadinle 2017.

Since our Gibbs procedure does not strictly enforce one-to-one matching,
it is possible for the final Bayes estimate to link multiple records in
\(B\) to one record in \(A\). The modeler can either report both such
matches (with their respective posterior match probabilities), or
resolve these conflicts by accepting only the match with highest
posterior probability. Such a resolution procedure indeed is equivalent
to minimizing posterior risk under the restriction of one-to-one
matching, and as thus theoretically justified within the Bayesian
framework.

\hypertarget{efficient-computation}{%
\section{Efficient Computation}\label{efficient-computation}}

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns. Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
that this bound does not scale with \(n_A\) or \(n_B\). Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
Wherever possible, we conduct calculations over these \(P\) agreement
patterns rather than the \(n_A \times n_B\) record pairs.

\hypertarget{data-representation-hashing-and-storage}{%
\subsection{Data Representation, Hashing, and
Storage}\label{data-representation-hashing-and-storage}}

The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
comparison vector as vector of length \(F\), with each component
\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). To ease
computations, we instead represent the comparison as a concatenation of
\(F\) many binary indicator vectors of lengths \(L_f\). For example, if
\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
under the classical framework becomes
\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
bijective transformation that does not change the meaning of the data,
but this representation will ease calculations and posterior updates.

In the classic Fellegi Sunter framework, \(\Gamma\) is a
\(n_A n_B \times F\) matrix, with each row providing the comparison
vector for a different \((i,j)\) pair. In contrast, we do not store
these comparison vectors themselves, but instead only a hashed value
corresponding to the agreement pattern of the \((i, j)\) pair. We store
this information in a nested list \(\tilde{\Gamma}\) where the
\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
each \(p\), we also calculate
\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
the total instances of agreement pattern \(p\) throughout the data, and
also for each \(j\), we calculate
\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
instances of agreement pattern \(p\) among the comparison vectors
between record \(j \in B\) and each of the \(n_A\) records in \(A\).

The hashing procedure described above considerably reduces the memory
needed to store the comparison information: instead of storing
\(n_A \times n_B\) comparison vectors, which are relatively long under
either the Fellegi Sunter or our modified framework, we only store the
\(P\) unique vectors, and then \(n_A \times n_B\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_A \times n_B\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

To address this, we propose a new strategy called \emph{storage
efficient indexing} (SEI). In standard indexing, the modeler decides a
certain criteria that they expect all true matching pairs to satisfy,
and a priori label any record pairs that do not meet that criteria as
non-matching. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. We
propose a method of reducing the comparison space (and reducing storage
requirements) without these drawbacks. Note that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose
to store only a small number \(R\) of them. Posterior calculations still
attribute the appropriate weight to all records through the summary
statistics \(H_p\), and \(H_{p_j}\). Rather than storing
\(n_A \times n_B\) record labels, SEI allows us to store at most
\(n_B \times P \times R\) labels, regardless of how large \(n_A\) is.

Laslty, for large data, we can partition the two datasets \(A\) and
\(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Thus the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

\hypertarget{gibbs-sampling}{%
\subsection{Gibbs Sampling}\label{gibbs-sampling}}

\underline{Updating $\mathbf{m}$ and $\mathbf{u}$:} After receiving
matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
field and level. This constitutes \(2 \times \sum L_f\) many summations
over \(n_A \times n_B\) quantities, and becomes computationally
burdensome with large data. In contrast, we recognize that each unique
agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
number of matching record pairs with agreement pattern \(p\), then the
contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
vector is simply \(H_p^m \times h_p\). Thus our posterior update for
\(\alpha\) is simply
\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
we can easily calculate \(H_p^u\), the number of nonmatching record
pairs of agreement pattern \(p\), by subtracting the number of matching
pairs from the total present in the data; that is
\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
parameter through
\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
that these constitute \(P\) many summations over \(n_B\) quantities, and
thus avoid the \(n_A \times n_B\) summation from the original method.

\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
full conditional provided earlier is conceptually straightforward, it
becomes computational burdensome when \(n_A\) is larger. The reader can
confirm that sampling a value from a large set of unequal probabilities
becomes difficult in most programming languages. To speed up
computation, we break this sampling step into two simpler steps. First,
we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
unique pattern and sample the agreement pattern between \(j\) and its
potential match. Second, we sample the record label uniformly among
records associated with that agreement pattern for that particular
\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
pattern between \(j\) and its potential match, and say
\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,

\[P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}\]
\[P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
\frac{1}{H_{p_j}} & (i, j) \in h_p \\
0 & \text{otherwise} \\
\end{cases}\]

Lastly, we recognize that all posterior updates are governed by the
agreement patterns of the record pairs rather than the record labels
themselves. Thus we complete the entire Gibbs procedure first at the
level of the \(P\) agreement patterns with the first equation above.
After, we can back-fill the records corresponding to the agreement
patterns through the second equation. Sampling uniformly is
computational simple even for large sets of candidate records, but this
step can also be parallelized when working with large data.

\hypertarget{simulation-study-1---accuracy}{%
\section{Simulation Study 1 -
Accuracy}\label{simulation-study-1---accuracy}}

In his 2017 paper, Sadinle demonstrated the accuracy of his method by
running \texttt{BRL} on simulated pairs of files with differing levels
of errors and overlap. Each file consists of first name, last name,
gender, postal code, age, and occupation, and following Sadinle's
implementation, we use only four fields for the linkage process: both
names, age, and occupation. Similarity scores for the names are binned
into three levels according the default settings for the
\texttt{compareRecords} function in the \texttt{BRL} package, and age
and occupation similarity scores are binary for exact matching. We use
flat priors for all \(\mathbf{m}\) and \(\mathbf{u}\) parameters, and
run the Gibbs sampler for 1000 iterations, with the first 100 discarded
as burn-in.

Comparing our method against \texttt{BRL} on these same simulated
datasets, we find that our method only has weakened performance in the
most extreme scenario of very high errors and very high overlap across
files. We use the following measures of accuracy in our analysis:
although using only recall and precision is common in the literature, we
find it useful to have a single metric to gauge overall performance, and
thus also include F measure.

\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot}

}

\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{simulation-study-2---speed}{%
\section{Simulation Study 2 - Speed}\label{simulation-study-2---speed}}

To demonstrate speed, we generate comparison vectors from pre-specified
distributions so that we can easily increase the size of the linkage
problem. Distributions are meant to emulate the behavior of similarity
scores across first name, last name, and day, month. For simplicity, we
consider only exact matching, so a vector (1, 0) corresponds to
agreeement and (0, 1) to disagreement. We simulate these data for
different values of \(n_A\) and \(n_B\), and compare the run-time of
\texttt{fabl} against \texttt{BRL}. Note that the number of unique
patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
consistently attained in the larger simulations.

We see that at low data size, \texttt{BRL} outperforms, but that
\texttt{fabl} is significantly faster at handling larger data. In
particular, run-time for \texttt{BRL} seems to grow quadratically (or
linearly with the size of both \(A\) and \(B\)) while run-time for
\texttt{fabl} seems to grow linearly (in the size of \(B\)).

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & m & u \\ 
  \hline
  fname & (0.95, 0.05) & (0.01, 0.99) \\ 
  lname & (0.95, 0.05) & (0.01, 0.99) \\ 
  day & (0.95, 0.05) & (0.03, 0.97) \\ 
  month & (0.95, 0.05) & (0.08, 0.92) \\ 
  year & (0.95, 0.05) & (0.07, 0.93) \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=17.75in]{../notes/figures/sadinle_speed_plot2}

}

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:unnamed-chunk-2}
\end{figure}

The above discussion suggests that for fixed \(n_B\), computation time
should remain mostly constant with growing \(n_A\). Simulation study
suggests that this is true. In the plot below, fixing \(n_B = 500\), we
see linear growth for the run-time under \texttt{BRL} as \(n_A\)
increases, with much more static run-time under \texttt{fabl}. The
slight increases in run-time that we do see are due primarily to the
hashing step, which again can be run in parallel for large data.

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB}

}

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:unnamed-chunk-3}
\end{figure}

We note here that \texttt{BRL} is coded in C, which makes for unfair
comparison against \texttt{fabl}, currently only built in R.
Additionally, although \texttt{fabl} is amenable to parallelization,
this simulation was run on a single core. Running \texttt{fabl} in C++
with paralellization for the hashing step and sampling the matching
status of the record pairs should lead to even more computational gains.

\hypertarget{data-analysis}{%
\section{Data Analysis}\label{data-analysis}}

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualities, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casuality lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

There are several challenges with working with such data. Firstly, both
datasets have been automatically digitized, which inherently leads to
some degree of typographical error. Secondly, the CDHES records are all
second hand accounts reported by individuals, which can result in
additional errors. Lastly, the only fields recorded are given name, last
name, date of death, and place of death; it is relatively common for a
parent and child to share the same given name, resulting in
indistinguishable records for two different individuals. This last point
nearly breaks the earlier mentioned assumption that there are no
duplicates within files, and reveals a key difference between
\texttt{BRL} and our proposed method.

We only utilized records with nonmissing entries for given and last
name, results in \(n_A = 4420\) files in CHDES and \(n_B = 1323\) files
in ERTL. The names were standardized to account for common misspellings
in the Spanish language, and then compared using a modified Levenstein
distance to account for the fact that second names are often ommited.
Place of birth is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, we excluded it from our linkage
process. Thus we conduct linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/m_posterior_smallP}

}

\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:unnamed-chunk-4}
\end{figure}

To mirror the original implementation, we constructed the comparison
vectors using 4 levels of agreement for each field, according to the
thresholds provided in Figure XX. This took 422 seconds for our proposed
method and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
aggreement levels creates unnecesary distictions in the data and makes
the comparison vectors less interpretable. Therefore we re-ran our
analysis with fewer agreement levels for each field (see Figure XX), and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method became much faster,
finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computional
gains can be made by simplifying the construction of the comparison
vectors. We also note that estimates of the \(\mathbf{m}\) and
\(\mathbf{u}\) paramters under each method are very similar, as shown in
Figure XX.

\hypertarget{violations-of-one-to-one-matching}{%
\subsection{Violations of One-to-One
Matching}\label{violations-of-one-to-one-matching}}

Figure XX shows the posterior distribution for \(D\), the number of
duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
see that \texttt{fabl} consistently overmatches within each Gibbs
iteration when compared to \texttt{BRL}, which is to be expected because
\texttt{BRL} explicitly prevents matches that violate one-to-one
matching throughout the entire sampler. Most of such matchings are due
to the randomness in the sampling procedure, and they occur sporadically
throughout the sampler in such a way that does not measurably influence
the eventual Bayes estimate.

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}

}

\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amonut of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:unnamed-chunk-5}
\end{figure}

Overall, both methods presented similar results. \texttt{fabl} yielded
an initial Bayes estimate of 195 matches found across files, and after
resolving matches that violated one-to-one requirements, yeilded 180
matches across files. This is acceptably close to the 181 matches found
by \texttt{BRL}. The main reason for this discrepancy is the difference
in how each model handles situations in which one file in \(B\) has
multiple plausible matches in \(A\).

\begin{figure}

{\centering \includegraphics[width=6.51in]{../notes/figures/el_salvador/rosa_records}

}

\caption{Example of linkage situation with multiple plausible matches}\label{fig:unnamed-chunk-6}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/bad_mixing}

}

\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:unnamed-chunk-7}
\end{figure}

The records shown in Figure XX provide one such instance. Note these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. Additionally, using the modified Levenstein distance
from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

Figure XX shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(B\) match to the
same record in \(A\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. Additionally,
\texttt{fabl} allows the modeller to inspect records with mulitple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

We have presented a method for Bayesian record linkage that is feasible
for large datasets. In particular, our hashing procedure and model
assumptions allow for a linkage procedure whose computational complexity
does not scale with the size of the larger dataset, making this method
particularly powerful in linking records when one datafile is
substantially smaller than the other.

In our case study, we included an exploration of how to conduct record
linkage when modelling assumptions are not met in practice. We explored
``one-to-many'' scenarios in which one record in \(A\) has multiple
plausible matches in \(B\), and showed how both \texttt{fabl} and
\texttt{BRL} demonstrated undesirable qualities. Other issues arise
under ``many-to-one'' scenarios, where one record in \(B\) has multiple
plausible matches in \(A\), and ``many-to-many'' scenarios in which
there is duplication both across and within datasets. Tuning
\texttt{fabl} for use in these scenarious is one potential avenue for
future work.

\end{document}