### 2021-09-22 Prelim/prelim.Rmd
added prelim slides

---
title: "Prelim"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Record Linkage

# Fellegi and Sunter

# Notation

- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors

- $Z_j$

- $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$

- $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$

- $\lambda = P(Z_j \leq n_A)$

# m's and u's

# Previous work

# Model

# Hashing

Demonstration

# Managing large data

# Simulation Results 1

# Simulation Results 2

# Extentions and Directions

- $m_{fl} = $

# Previous work

______________________________________

# El Salvador

# Distributed computation

# Storage Efficient Indexing