### 2021-09-23 Prelim/prelim.Rmd
updated prelim slides

---
title: "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
```

# Introduction to Record Linkage

<!-- \begin{columns} -->
<!-- \begin{column}{0.5\textwidth} -->

<!-- \begin{itemize} -->
<!-- \item Record linkage is the task of identifying duplicate records over noisy datasets. -->

<!-- \item Easy with unique identifiers, difficult when faced with errors -->

<!-- \item Far ranging applications in business, public health, and human rights -->

<!-- \end{itemize} -->

<!-- \end{column} -->
<!-- \begin{column}{0.5\textwidth}  %%<--- here -->
<!--     \begin{center} -->
<!--      picture -->
<!--      \end{center} -->
<!-- \end{column} -->
<!-- \end{columns} -->

- Record linkage is the task of identifying duplicate records over noisy datasets.

- Easy with unique identifiers, difficult when faced with errors

- Far ranging applications in business, public health, and human rights

# Examples

```{r}
knitr::include_graphics("rl_examples.png")
```

# Fellegi and Sunter

```{r out.width="100%"}
knitr::include_graphics("fs_graphic.png")
```
# Previous work

- Fellegi and Sunter (1969)
  - independent classification of all $n_A n_B$ record pairs. Transitive closure acheived through postprocessing

- Enamorado et al (2019): \texttt{fastlink} 
  - provides efficient and scalable FS model

- Sadinle (2017) - Beta Record Linkage (\texttt{BRL}) 
  - Bayesian method for bipartite matching, strictly enforces one-to-one matching. Does not scale well to large linkage tasks

# Notation

- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$

- $Z_j = \begin{cases} 
	i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
	n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
	\end{cases}$

- $m_{fl} = P\left(\gamma_{ij}^f = l |Z_j = i\right)$

- $u_{fl} = P\left(\gamma_{ij}^f = l |Z_j \neq i\right)$

- $\lambda = P(Z_j \leq n_A)$

# Fast Beta Linkage (fabl)

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$

$$\mathbf{m_{f}} \sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{fL_f})$$
$$\mathbf{u_{f}} \sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{fL_f})$$
$$Z_j | \lambda =
\begin{cases} 
	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
	1-\lambda &  z_j  = n_A + 1 \\
\end{cases}$$
$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$

- Model specification allows for **parallel/distributed** computing, **hashing** of comparison vectors, and **storage efficient indexing (SEI)**

# Hashing

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $h(i, j)=p.$

<!-- (i, j) \in h_p$. -->

- Reduce data to sufficient statistics
  - $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
  - $H_{p_j} = ||r_{p_j}||$
  - $H_p = \sum_{j} H_{p_j}$

- Run Gibbs sampler at level of agreement patterns, not record pairs
  - Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
  - Use number of matches for each pattern to update $m$ and $u$
  - Back fill record labels at the end through $r_{p_j}$
  - Computational complexity $O(P \times n_B \times F)$ instead of $O(n_A \times n_B \times F)$

<!-- - Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$ -->

# Managing Large Data

- **Distributed Computing** - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results

- **Storage Efficient Indexing (SEI)** - Store at most small number $R$ many record labels in each $r_{p_j}$, remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$.

- Hashing plus SEI can reduce memory requirements by >99%.
  - Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.

# Three Simulations

- We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
  - Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
  - Measure speed when both $n_A$ and $n_B$ are increasing
  - Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed.

# Simulation Results 1

```{r}
knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
```

# Simulation Results 2

```{r}
knitr::include_graphics("../notes/figures/sadinle_speed_plot2.png")
```

# Simulation Results 3

```{r}
knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
```

# Extentions and Directions

- Linkage in the face of duplicates within datasets

- Models that allow reliability of information to differ by subgroup in the data

- Linkage over blocked data (allows for much larger linkage tasks)

# Questions?

# Additional Slides

- Details of intuition behind $m$ and $u$ parameters

- Loss function and Bayes estimate

- Details of one-to-one matching; conflict resolution under \texttt{fabl};

- El Salvador case study (time under different values of $P$, violations of one-to-one matching)

- Further details on distributed computation of comparison vectors and storage efficient indexing