### 2021-10-11 methods_paper/fabl-paper.tex
edited Efficiency section

\documentclass[12pt,letterpaper]{article}

\usepackage[hypertexnames=false]{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}

% Custom packages
\usepackage{amsthm, amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tabularx}
\usepackage{tikz}           % vector graphics
\usepackage{bm}             % bold math symbols
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{import}
\usepackage{titling}
\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
  calc, shadows, plotmarks}

\graphicspath{{./notes/figures/}}

% Custom commands
\let\oldvec\vec
\renewcommand\vec{\bm}
\newcommand{\simfn}{\mathtt{sim}} % similarity function
\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
\newcommand{\distfn}{\mathtt{dist}} % distance function
\newcommand{\valset}{\mathcal{V}} % attribute value set
\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
\newcommand{\euler}{\mathrm{e}} % Euler's constant
\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
\def\spacingset#1{\renewcommand{\baselinestretch}%
  {#1}\small\normalsize} \spacingset{1}

\newtheorem*{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem*{definition}{Definition}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\sectionfont{\large\nohang\centering\MakeUppercase}

\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts \textsuperscript{b}}
\date{
 \textsuperscript{a} Department of Statistical Science, Duke University \\
 \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
  \today}

\begin{document}
\maketitle

\bigskip
\begin{abstract}
%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing. In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity. To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.

In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity. In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy. Finally, we discuss our work and future directions.

\end{abstract}

\noindent%
{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing

\newpage
\spacingset{1.5}

\section{Introduction}
\label{sec:introduction}

%	I have some proposed revisions for the first part of the paper. In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale. 
%	
%	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
%	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction. I've seen this done in many papers. 
%	
%	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
%	
%	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

	Many statistical record linkage methods are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.

	The independent pairwise matching assumption from Fellegi and Sunter is popular mainly for its mathematical simplicity, but is often unreasonable in practice. In many situations, we know that there are no duplications within a database, meaning that one record from one database should be linked with at most one record in the other database. Here, additional declared matches are known by assumption to be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of permitted matches under the one-to-one constraint, but the method was so computationally intensive it could used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (BRL), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.

	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference. Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, in Sections \ref{sec:simulations} and \ref{sec:case-studies} we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}. We conclude in Section \ref{discussion} with a discussion of directions for future work.

	% I think this is clear and has already been said. 
%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')

%

\section{Review of Prior Work}
\label{sec:review-of_prior-work}

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.

\subsection{Comparison Data}

Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

%In what follows we use notation that reflects that these comparison vectors can be modelled in number of ways. In practice however, this binning in beneficial because it allows us to model any type of comparison through a multinomial distribution. In a Bayesian context, Dirichlet priors for

\subsection{Fellegi-Sunter Model}
The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.

In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume that comparison vectors are independent given their matching status, and that match status of the record pairs are independent, allowing us to model 
the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs. These unknown parameters are often estimated using the EM algorithm.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined). This is a problem in practice as transitive closures are violated. Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
\begin{align}
	\label{eqn:jaro}
	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_1. \notag
\end{align}

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. \cite{jaro1989} provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate.

%One criticism of using mixture models as proposed by \cite{jaro1989} is the fact that the decision rule will lead to ``many-to-many assignments (breaking transitive closures). One way of resolving this issue to incorporate the constraint directly into the model \citep{Fortinietal01, Matsakis10, liseo_jos_2011, liseo_2011, Larsen05, Larsen12, gutman_bayesian_2013, sadinle_bayesian_2017}, which is appealing as one then obtains uncertainty quantification under the Bayesian paradigm, but often has to sacrifice scaling to very large databases. On the other hand, one can turn to post-processing methods in practical situations. In this paper, we attempt to bridge the best of both approaches. Specifically, we propose a model in the spirit of \cite{sadinle_bayesian_2017} and \cite{enamorado_using_2019}. We relax the one-to-one matching constraint proposed by  \cite{sadinle_bayesian_2017} for purely computational reasons, and resolve this in a post-processing step. In addition, we propose our own computational speeds up and those in the spirit of \cite{enamorado_using_2019}. Our framework is then embarassingly parallel, available in the \texttt{fabl} package, and evaluated on a real case study and simulation studies illustrating that any loss in accuracy due to removing the constraint is minimal.

\subsection{Beta Record Linkage Model}

One criticism of using mixture models is that the decision rule will lead to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

\subsection{Blocking}

In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or ``blocks'' to reduces this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, the method the follows can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.

\section{Fast Beta Linkage}
\label{sec:fast-beta-linkage}

Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink. In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we propose independent priors for each component $Z_j$. We denote the \emph{fast beta prior} as follows:

$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + j \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$.. One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\mathbf{Z}\). We encourage the reader to integrate out $\pi$ from the $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.

Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimate that proportion of record pairs that have matches. We find $\pi$ to be more interpretable parameter than $\lambda$, especially considering that the overwhelming majority of pairs are non matches, meaning that $\lambda$ necessarily tends towards 0 as the size of the linkage task grows. Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.

Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

%this independent prior for each \(Z_j\), induces a sampler that is \emph{embarassingly parallel}, allowing for significant computational gains over the Gibbs sampler under \texttt{BRL}. More importantly, since only the agreement pattern of \(Z_j\) is used forcalculations within the Gibbs sampler, and not the particular record label, this sampling is only needed at the level of the unique agreement patterns, which offers additional computational savings.

\hypertarget{fabl-model}{%
	\subsection{Model Specification}
	\label{model-specification}}

As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent. Additionally, we assume that any missing data is missing at random. Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches. In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.
The model can then be specified as follows:

$$\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) = \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}$$
$$\bm{m}_f \sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f})$$
$$\bm{u}_f \sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}).$$
$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + j \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$$

\hypertarget{posterior-sampling}{%
	\subsection{Gibbs Sampler}
	\label{gibbs_sampling}}
We work with the following factorization of the joint posterior distribution:

\begin{align*}
	p(\mathbf{Z}, \Phi, \pi|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \Gamma^{obs}) p(\mathbf{Z} | \pi) p(\Phi) p(\pi) \\
	&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
	&\times \prod_{j=1}^{n_2} \left[I(Z_j \leq n_1)\frac{1}{n_1}\pi + I(Z_j > n_1)(1 - \pi)\right] \\
	&\times \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1}
\end{align*}
This factorization leads to following Gibbs Sampler:

\underline{Sample $\Phi^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
standard multinomial-dirichlet mechanics. Thus we have

\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)\)
and
\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)\).

\underline{Sample $\pi^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
\(\pi\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\pi) = \mathcal{L}(\pi|\mathbf{Z})\) is
determined only by the number of duplicates
$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
$\mathbf{Z}$. Thus we have

\begin{align*}
p(\pi | \mathbf{Z}) &\propto p(\mathbf{Z}|\pi)p(\pi) \\
&\propto \pi^{n_{12}(\mathbf{Z})} (1-\pi)^{n_2 - (n_{12}(\mathbf{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
&\propto \pi^{n_{12}(\mathbf{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\pi} -1} \\
\implies &\pi^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi})
\end{align*}

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \Phi^{(s+1)}, \pi^{(s+1)}$:}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\), and note that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\)
is a discrete distribution with probabilities proportional to

\begin{align*}
	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&=
	\begin{cases} 
		w_{ij}  & z_j \leq n_1; \\
		1 &  z_j  = n_1 + j \\
	\end{cases}\\
\end{align*}

where
$w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}$.
The interested reader should note that these are precisely the
likelihood ratios used in the Fellegi-Sunter model to classify matches
and non-matches, and we therefore refer to \(w_{ij}\) as the
\emph{Fellegi Sunter weights}.

With the likelihood in this form, we can derive the full conditional
\[p(Z_j|\Gamma_{.j}, \Phi, \pi) \propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi)\]

\[\propto \left(\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1}\right)\]
\[= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\mathbf{1}_{z_j = i} + (1-\pi)\mathbf{1}_{z_j = n_1 + 1} \]
\[ \implies Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto
\begin{cases} 
	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + 1 \\
\end{cases}\]

In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
integrate over the posterior of \(\pi\) and rearrange terms to
produce the final full conditional:

\[p\left(Z_j^{(s+1)}  = i| \Phi, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{ij}  & i \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\mathbf{Z}) + \beta_{\pi}}{n_{12}(\mathbf{Z}) + \alpha_{\pi}} & i  = n_1 + 1 \\
\end{cases}\]

\hypertarget{bayes-estimate}{%
	\subsection{Bayes Estimate}
	\label{bayes-estimate}}

We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
parameter \(\mathbf{Z}\) by assigning different positive losses to
different types of errors, and minimizing posterior expected loss. We
adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
option to leave the matching undetermined by the model. Specifically, we have

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model.

We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$. In this paper, we adopt losses
\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule

\[\hat{Z}_j =\begin{cases} 
	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
	0,  & \text{otherwise} ; \\
\end{cases}\]

For a more in-depth explanation of this function and the induced Bayes
estimate, see \citep{sadinle_bayesian_2017}.

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study.

To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.

%\newpage
%
%\hypertarget{posterior-sampling}{%
%	\subsection{Gibbs Sampler}
%	\label{gibbs_sampling}}
%We work with the following factorization of the joint posterior distribution:
%
%\textcolor{red}{Brian: The Gibbs sampler has some minor typos. It should be written in terms of $\bm{Z}, \Phi, \lambda \mid \gamma_{obs}.$ Also, just make the rest of the notation match please:)}
%
%\begin{align*}
%p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) &\propto \mathcal{L}(\bm{Z}, \Phi \mid \bm{\gamma}^{obs}) p(\mathbf{Z} | \lambda) p(\Phi) p(\lambda) \\
%&\propto \prod_{j=1}^{n_2}  \prod_{i=1}^{n_1}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)} \\
%&\times  \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{\alpha_{fl} - 1}  \times\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{\beta_{fl} - 1} \\
%&\times \prod_{j=1}^{n_B} \left[I(Z_j \leq n_1)\frac{1}{n_1}\lambda + I(Z_j > n_1)(1 - \lambda)\right]
%\end{align*}
%This factorization leads to following Gibbs Sampler:
%
%\underline{Sample $\mathbf{m}^{(s+1)}$ $\mathbf{u}^{(s+1)}|\Gamma, \mathbf{Z}^{(s)}$:}
%The \(\mathbf{m}\) and \(\mathbf{u}\) parameters are updated through
%standard multinomial-dirichlet mechanics. Thus we have
%
%\[\mathbf{m}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\mathbf{Z}), \ldots, \alpha_{fL_f}(\mathbf{Z}))\]
%\[\mathbf{u}_f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\mathbf{Z}), \ldots, \beta_{fL_f}(\mathbf{Z}))\]
%where
%\(\alpha_{fl}(\mathbf{Z})= \alpha_{fl} + \sum_{i,j} I(\gamma_{ij}^f = l) I(z_j = i)\)
%and
%\(\beta_{fl}(\mathbf{Z})= \beta_{fl} + I(\gamma_{ij}^f = l) I(z_j \neq i)\).
%
%\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
%\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
%of successes (when \(z_j < n_A + 1\)) and failures (when
%\(z_j = n_A + 1\)), and therefore
%\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
%determined only by the number of duplicates
%$n_{12}(\mathbf{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by
%$\mathbf{Z}$. Thus we have
%
%\[p(\lambda | \mathbf{Z}) \propto p(\mathbf{Z}|\lambda)p(\lambda)\]
%\[\propto \lambda^{n_{12}(\mathbf{Z})} (1-\lambda)^{n_2 - (n_{12}(\mathbf{Z})} \lambda^{\alpha_{\lambda} -1} (1-\lambda)^{\beta_{\lambda} -1}\]
%\[ \propto \lambda^{n_{12}(\mathbf{Z}) + \alpha_{\lambda} - 1} (1-\lambda)^{n_1 - n_{12}(\mathbf{Z}) + \beta_{\lambda} -1}\]
%\[\implies \lambda^{(s+1)}|\mathbf{Z}^{(s+1)} \sim \text{Beta}(D + \alpha_{\lambda}, n_2 - n_{12}(\mathbf{Z}) + \beta_{\lambda})\]
%
%\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
%Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
%only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
%denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
%that as a function of \(Z_j\), the likelihood
%\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
%is a discrete distribution with probabilities proportional to
%
%\begin{align*}
%	p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}\\
%	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i, \gamma_{ij}^f = l)} \\
%	&=
%	\begin{cases} 
%		w_{ij}  & z_j \leq n_1; \\
%		1 &  z_j  = n_1 + j \\
%	\end{cases}\\
%\end{align*}
%
%where
%\(w_{ij} = \left(\frac{\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}}\right)^{I(\gamma_{ij}^f = l)} = \frac{P(\boldsymbol{\gamma_{ij}}|Z_j = i)}{P(\boldsymbol{\gamma_{ij}} |Z_j \neq i)}\).
%The interested reader should note that these are precisely the
%likelihood ratios used in the Fellegi-Sunter model to classify matches
%and non-matches, and we therefore refer to \(w_{ij}\) as the
%\emph{Fellegi Sunter weights}.
%
%With the likelihood in this form, we can derive the full conditional
%\[p(Z_j|\Gamma_{.j}, \mathbf{m} ,\mathbf{u}, \lambda) \propto p(\Gamma_{.j}| Z_j, \mathbf{m} ,\mathbf{u}) P(Z_j|\lambda)\]
%
%\[\propto \left(\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + \mathbf{1}_{z_j = n_A + 1}\right)\left(\lambda\sum_{i=1}^{n_A}\frac{1}{n_A}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1}\right)\]
%\[= \frac{\lambda}{n_A}\sum_{i=1}^{n_A}w_{ij}\mathbf{1}_{z_j = i} + (1-\lambda)\mathbf{1}_{z_j = n_A + 1} \]
%\[ \implies Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \Gamma, \lambda \propto
%\begin{cases} 
%	\frac{\lambda}{n_A}w_{ij}   & z_j \leq n_A; \\
%	1-\lambda &  z_j  = n_A + 1 \\
%\end{cases}\]
%
%In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
%integrate over the posterior of \(\lambda\) and rearrange terms to
%produce the final full conditional:
%
%\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
%\begin{cases} 
%	w_{ij}  & i \leq n_A; \\
%	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
%\end{cases}\]
%
%\hypertarget{bayes-estimate}{%
%	\subsection{Bayes Estimate}
%	\label{bayes-estimate}}
%
%We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
%parameter \(\mathbf{Z}\) by assigning different positive losses to
%different types of errors, and minimizing posterior expected loss. We
%adopt the loss function proposed in Sadinle 2017, in which
%\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
%option to leave the matching undetermined by the model. Specifically, we have
%
%\[L(\hat{Z_j}, Z_j)=\begin{cases} 
%	0  & \text{if } Z_j = \hat{Z_j}; \\
%	\theta_R,  & \text{if } \hat{Z_j} = R; \\
%	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
%	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
%	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
%\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
%on the linkage status, \(\theta_{10}\) is the loss from a false
%non-match, \(\theta_{01}\) is the loss from a false match, and
%\(\theta_{11}\) is the loss from the special case of a false match in
%which the record has a true match other than the one estimated by the
%model. 
%
%We seek to minimize the posterior expected loss, given by $\mathbb{E}[L(\hat{\mathbf{Z}}, \mathbf{Z})|\Gamma^{obs}] = \sum_{\mathbf{Z}} L(\hat{Z_j}, Z_j)P(\mathbf{Z}|\Gamma^{obs})$. In this paper, we adopt losses
%\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\), inducing the intuitive decision rule
%
%\[\hat{Z}_j =\begin{cases} 
%	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
%	0,  & \text{otherwise} ; \\
%\end{cases}\]
%
%For a more in-depth explanation of this function and the induced Bayes
%estimate, see \citep{sadinle_bayesian_2017}.
%
%Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study. 
%
%To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that two matches. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.

\section{Efficient and Scalable Implementation}
\label{sec:efficiency}

To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we employ hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method. The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$. Here, 
$P = |\mathcal{P}|$ is the total number of unique agreement
patterns; observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\). In this context, the integers $1, \ldots, P$ serve as \emph{hashed values} that encode the same information as the orignial vectors themselves. Whenever possible, we conduct calculations over these \(P\) agreement
patterns, instead of the typical \(n_1 \times n_2\) record pairs.

Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations.

%In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model. 
% Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
%through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
%\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
%\(P = |\mathcal{H}|\) denote the total number of unique agreement
%patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
%that this bound does not depend on \(n_1\) or \(n_2\). Prior to
%processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
%and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
%Next, we map record pairs to their corresponding hashed value. 
%More specifically, when the record pair \((i,j)\) 
%exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
%Whenever possible, we conduct calculations over these \(P\) agreement
%patterns, instead of the typical  \(n_1 \times n_2\) record pairs.

\hypertarget{data-representation-hashing-and-storage}{%
	\subsection{Data Representation, Hashing, and
		Storage}\label{data-representation-hashing-and-storage}}

First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019using} accomplished this efficiently through the hashing function

$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$

This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair. For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns in $\mathcal{P}$. When the $(i, j)$ pair exhibits the $p^{th}$ pattern, we say $(i, j) \in h_p$.

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics. Note that $\mathcal{P}$ and $\mathcal{R}$ fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}\}$ for our efficient posterior inference.

\hypertarget{large-data}{%
	\subsection{Working with Large Data}\label{large-data}}

The hashing procedure described above considerably reduces the memory
needed to store the comparison information. That is, instead of storing
\(n_1 \times n_2\) comparison vectors, we only store the
\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_1 \times n_2\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

%The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
%comparison as a vector of length \(F\), with each component
%\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
%To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
%\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
%under the classical framework becomes
%\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
%bijective transformation that does not change the meaning of the data,
%but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.

%In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
%\(n_1 n_2 \times F\) matrix, with each row providing the comparison
%vector for a different \((i,j)\) pair. In contrast, we do not store
%these comparison vectors themselves, but instead only the hashed value $h_p$
%corresponding to the agreement pattern of the \((i, j)\) pair. We store
%this information in a nested list $\tilde{\bm{\Gamma}}$ where the
%\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
%in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
%each \(p\), we also calculate
%\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
%the total instances of agreement pattern \(p\) throughout the data, and
%also for each \(j\), we calculate
%\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
%instances of agreement pattern \(p\) among the comparison vectors
%between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).

To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI). In standard indexing, one decides a priori certain criteria that they expect all true matching pairs to satisfy, 
and labels any record pairs that do not meet that criteria as
non-matches. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error.

With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\). Therefore we know that records $\i in r_{j_p}$  such that
\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Posterior calculations still attribute the appropriate weight to all records through the summary statistics \(H_p\), and \(H_{j_p}\). Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.

Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.  Thus, the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

\hypertarget{efficient-posterior}{%
	\subsection{Efficient Posterior Inference }\label{efficient-posterior}}

\underline{Sample $\Phi^{(s+1)}|\tilde{\Gamma}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities. These are simple calculations, but become computationally burdensome when working on large linkage tasks.

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote
$H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the
number of matching record pairs with agreement pattern $p$. We can also easily calculate the number of nonmatching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\). Then, if $\mathbf{\alpha_0}$ and $\mathbf{\beta_0}$ are vectors of prior parameters for the $\mathbf{m}$ and $\mathbf{u}$ distributions respectively, the posterior update becomes simply

$$\alpha(\mathbf{Z}) = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
$$\beta(\mathbf{Z}) = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$

Note that these constitute \(P\) many summations over \(n_B\) quantities, and thus avoid the \(n_A \times n_B\) summation from the original method.

%After receiving
%matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
%\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
%field and level. This constitutes \(2 \times \sum L_f\) many summations
%over \(n_A \times n_B\) quantities, and becomes computationally
%burdensome with large data. In contrast, we recognize that each unique
%agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
%and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
%\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
%number of matching record pairs with agreement pattern \(p\), then the
%contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
%vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
%\(\alpha\) vector is simply
%\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
%we can easily calculate \(H_p^u\), the number of nonmatching record
%pairs of agreement pattern \(p\), by subtracting the number of matching
%pairs from the total present in the data; that is
%\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
%parameter through
%\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
%that these constitute \(P\) many summations over \(n_B\) quantities, and
%thus avoid the \(n_A \times n_B\) summation from the original method.

\underline{Sample $\mathbf{Z}^{(s+1)}|\tilde{\Gamma}, \Phi^{(s+1)}$:} Although sampling \(Z_j\) from a the
full conditional provided earlier is conceptually straightforward, it
becomes computational burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up
computation, we break this sampling step into two simpler steps. First,
we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
unique pattern and sample the agreement pattern between \(j\) and its
potential match. Second, we sample the record label uniformly among
records associated with that agreement pattern for that particular
\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
pattern between \(j\) and its potential match, and say
\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,

$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{p}\times H_{j_p}  & p \leq P; \\
	n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}$$
$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) \begin{cases} 
	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
	0 & \text{otherwise} \\
\end{cases}$$

Lastly, we recognize that all posterior updates are governed by the
agreement patterns of the record pairs rather than the record labels
themselves. Thus we complete the entire Gibbs procedure first at the
level of the \(P\) agreement patterns with the first equation above.
After all Gibbs iterations are complete, we can back-fill the records corresponding to the agreement patterns through the second equation.

We emphasize the computational gains of this split sampler: the first step is a sample from $P$ many options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records. These changes in themselves drastically improves the speed of the sampler, and each can be parallelized if desired for additional computational gains.

To aid the reader, we provide summary of the \texttt{fabl} method through pseudocode:

\begin{algorithm}[h!]
	\caption{Summary of fabl algorithm}
	\begin{algorithmic}[1]

		\Procedure{Hashing and Preprocessing}{}
		\State Partition files $A$ and $B$ into chunks $\{A_I\}, \{B_J\}$
		\For{each $I$, $J$}
		\State Create comparison vectors between $A_I$ and $B_J$
		\State Hash results and calculate summary statistics
		\State Use SEI to reduce memory usage
		\EndFor
		\State Synthesize results across pairings
		\EndProcedure

		\Procedure{Gibbs Sampling}{}
		\State Initialize $m$, $u$, and $Z$ parameters

		\For{$t \in \{1, \ldots, T\}$} 
		\State Sample $m^{t+1}|Z^{t}, \Gamma$ and  $u^{t+1}|Z^{t}, \Gamma$ 
		\State Sample $H\left(Z^{t+1}\right)|m^{t+1}, u^{t+1}, \Gamma$  \Comment{Sample agreement pattern, not record}
		\EndFor
		\State Sample $Z | H(Z), \Gamma$ \Comment{Fills in record label based on agreement pattern}

		\EndProcedure

	\end{algorithmic}
\end{algorithm}

\hypertarget{computational-complexity}{%
	\subsection{Computational Complexity}\label{computational-complexity}}

{
\color{blue}
The computational complexity of \texttt{fabl} is determined by its two steps: constructing comparison vectors and Gibbs sampling. The computational complexity of all pairwise comparisons across files $\mathbf{X}_1$ and $\mathbf{X}_2$ is $O(F n_1 n_2)$. Since \texttt{fabl} is able to compute these comparisons in parallel, we can split these computations across $B$ equally sized partitions, so the complexity becomes $O(\frac{F}{B} n_1 n_2)$.

Without hashing, the computational complexity of updating the $\mathbf{m}$ and $\mathbf{u}$ parameters is $O(F n_1 n_2)$. However, by doing calculations over the agreement patterns rather than the individual record, hashing reduces the complexity to $O(FP)$.

The complexity of updating $Z$ sequentially at the record level is $O(n_1 n_2)$. With hashing, we split this sampling into two steps; first we sample that agreement pattern of the match with complexity $O(n_2 P)$, and then we sample the record that exhibits that pattern with complexity $O(n_2)$. Thus we have reduced the complexity of the Gibbs sampler under \texttt{BRL} from $O(F n_1 n_2)$ to $O(n_2 P)$.

Overall, the computational complexity under the complete \texttt{fabl} framework is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$.
}

\section{Simulation Studies}
\label{sec:simulations}

In this section, we replicate a simulation study from \cite{sadinle_bayesian_2017} to compare \texttt{fabl} against \texttt{BRL} on several simulated datasets with varying amounts of error and overlap. We use first name, last name, age, and occupation for this linkage, and create comparison vectors according to the default settings of the \texttt{compareRecords} function from the \texttt{BRL} package. See Table~\ref{Tab:sadinle_simulation_cutoffs} for details on the construction of comparison vectors. Each simulation identifies duplicated individuals between two datasets, each with 500 records. We conduct linkage when matching records exhibit 1, 2, and 3 errors across the four fields, and when there are 50, 250, and 450 individuals in common across datasets. We use flat priors for all $m$ and $u$ parameters, run the Gibbs Sampler for 1000 iterations, and discard the first 100 as burn-in.

\begin{table}[h!]
	\centering
	\begin{tabular}[t]{ll|llll}
		\hline
		\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
		\cline{3-6}
		Fields & Similarity & 1 & 2 & 3 & 4\\
		\hline
		First and Last Name & Levenstein & 0 & (0, .25] & (.25, .5] & (5, .1]\\
		\hline
		Age and Occupation & Binary & Agree & Disagree &  & \\
		\hline
	\end{tabular}
\caption{Construction of comparison vectors for accuracy study with simulated datasets}
\label{Tab:sadinle_simulation_cutoffs}
\end{table}

\hypertarget{accuracy}{%
	\subsection{Precision, Recall, and F-measure}\label{accuracy}}

In this section, we compare \texttt{fabl} to 	\texttt{BRL} in terms of precision, recall, and f-measure, which are commonly defined record linkage evaluation metrics that are defined in \citep{christen_2012}. 
%
In cases when there are only one or two errors in matching records, and in cases with low to moderate duplication across records, we see that \texttt{fabl} provides approximately equivalent accuracy as \texttt{BRL}. 
We find that our method only has weakened performance in the
most extreme scenario of very high errors and very high overlap across
files. In these situations, \texttt{BRL} is removing large numbers of records from consideration throughout the Gibbs Sampler, making its implementation most different from \texttt{fabl}. However,  such extreme linkage tasks, which such high amounts of errors and overlap, are extremely rare in practice.

%\begin{figure}
%	\[\text{Recall} = \frac{\text{Matches Correctly Identified}}{\text{True Matches in Data}}\]
%	\[\text{Precision} = \frac{\text{Declared Matches}}{\text{True Matches in Data}}\]
%	\[\text{F-Measure} = 2\left(\frac{\text{Recall} \times \text{Precision}}{\text{Recall} + \text{Precision}}\right)\]
%	
%	\caption{Definitions of performance metrics}\label{fig:performance_metrics}
%\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot2} 
\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from \cite{sadinle_bayesian_2017}. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}
\label{fig:sadinle_simulation}
\end{center}
\end{figure}

%
%\begin{figure}[ht]
%	
%	{\centering 
%		
%		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot} 
%	}
%	
%	\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:sadinle_simulation}
%	
%\end{figure}

\hypertarget{partial}{%
	\subsection{Accuracy under Partial Estimates}\label{partial}}

Lastly, we review the performance of \texttt{fabl} and \texttt{BRL} using loss functions that allow partial estimates of the linkage parameter. By leaving $\theta_{10} = \theta_{01} = 1$ and $\theta_{11} = 2$, but setting $\theta_R = 0.1$, we allow the model to decline to decide a match for certain records, with nonassignment being 10\% as costly as a false match. In this context, we are no longer focused on finding all true matches, but rather protecting against false matches. Thus, instead of recall, we use the \emph{negative predictive value} (NPV), defined as the proportion of non-links that are actual non-matches. Mathematically, $NPV = \sum_{j=1}^{n_2} I(\hat{Z}_j = Z_j = n_1 + j)$/$\sum_{j=1}^{n_2} I(\hat{Z}_j = n_1 + j)$. We continue to use the precision, which is renamed the \emph{positive predictive value} in this context. Lastly, we also report that rejection rate (RR), or how often the model declines to make a linkage decision, defined as RR = $\sum_{j=1}^{n_2} I(\hat{Z}_j = R)$.

We see that \texttt{fabl} maintains equivalently strong PPV as \texttt{BRL} across all linkage settings. However, with high amounts of overlap and high amounts of error, the rejection rate under \texttt{fabl} rises, leading to a decrease in NPV. This is reasonable because large numbers of errors lead to situations where one record in $\bf{X}_2$ has multiple plausible matches in $\bf{X}_1$, so that posterior match probability is split across more records. Since \texttt{BRL} removes previously matched records from the set of available records for a new match, it is able to retain higher confidence in matches. In practice, this partial estimate approach is not often used, so we remain confident in \texttt{fabl}'s performance. If one wishes to use partial estimates, \texttt{fabl} will possibly leave more linkages for the modeller to match by hand that would be left under \texttt{BRL}, but the decisions made by each method will have nearly equal accuracy.

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial} 
	\caption{Negative predictive value (NPV) and positive predictive value (PPV) on simulated datasets}
	\label{fig:sadinle_simulation_partial}
\end{center}
\end{figure}

\hypertarget{speed}{%
	\subsection{Speed}\label{speed}}

To demonstrate speed, we generate comparison vectors from pre-specified
distributions so that we can easily increase the size of the linkage
problem. Distributions are meant to emulate the behavior of similarity
scores across first name, last name, and day, month. For example, $u^{\text{month, 1}} = P(\text{Records have same birth-month | Nonmatch}) = \frac{1}{12}$. For simplicity, we
consider only exact matching, so a vector (1, 0) corresponds to
agreement and (0, 1) to disagreement. We simulate these data for
different values of \(n_A\) and \(n_B\), and compare the run-time of
\texttt{fabl} against \texttt{BRL}. Note that the number of unique
patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
consistently attained in the larger simulations.

We see that at low data size, \texttt{BRL} outperforms, but that
\texttt{fabl} is significantly faster at handling larger data. In
particular, run-time for \texttt{BRL} seems to grow quadratically (or
linearly with the size of both \(A\) and \(B\)) while run-time for
\texttt{fabl} seems to grow linearly (in the size of \(B\)).

\begin{table}[h!]
	\centering
	\begin{tabular}{rrr}
		\hline
		& m & u \\ 
		\hline
		fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
		day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
		month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
		\hline
	\end{tabular}
\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
\end{table}

\begin{figure}[h!]

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2}

	}

	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$. We see near quadratic growth in runtime for BRL, and near linear growth for fabl.}\label{fig:speed1}
\end{figure}

The above discussion suggests that for fixed \(n_B\), computation time
should remain mostly constant with growing \(n_A\). Simulation study
suggests that this is true. In the plot below, fixing \(n_B = 500\), we
see linear growth for the run-time under \texttt{BRL} as \(n_A\)
increases, with much more static run-time under \texttt{fabl}. The
slight increases in run-time that we do see are due primarily to the
hashing step, which again can be run in parallel for large data.

\begin{figure}[h!]

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB}

	}

	\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500. We see linear growth in runtime for BRL, and near constant runtime for fabl.}\label{fig:speed2}
\end{figure}

We note here that \texttt{BRL} is coded in C, which makes for unfair
comparison against \texttt{fabl}, currently only built in R.
Additionally, although \texttt{fabl} is amenable to parallelization,
this simulation was run on a single core. Running \texttt{fabl} in C++
with paralellization for the hashing step and sampling the matching
status of the record pairs should lead to even more computational gains.

\hypertarget{scale}{%
	\subsection{Scaling to a Larger Simulation Study}\label{scale}}

In this section, we investigate scaling to larger data sets using our proposed methodology.

Specifically, we demonstrate how partitioning and hashing the data through \texttt{fabl} method allows us undertake significantly larger linkage tasks. We concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{} to create two larger files, each of 20,000 records (40,000 records total). Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting a final comparison matrix about 6.4 GB in size. Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task. Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.

For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact. We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons. The resulting data object is now only 90 MB, about 1\% the size of the object required under the standard method. Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run. However, this could be substantially sped up using distributed computing.

Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure. The reader will note that this slightly worse performance than witnessed in the smaller simulation studies; this is expected because it is naturally more difficult to link more files with the same amount of information. With more linkage fields \texttt{fabl} maintains the high accuracy seen above.

%\textcolor{blue}{This is with post processing already. All I meant here was that is that small tasks require less information. If you do a $500 \times 500$ linkage task and you only have first and last name (for example), thats likely enough to get all the correct matches. However, once you do $20000 \times 20000$, its more likely that more than one person in the second file has the same first and last name, so you need more information to distinguish. I mentioned before that the simulated files actually have 6 fields of information, but I only used 4 originally because thats what Sadinle did. I can use all 6 though to see what happens?}

\section{Application to Civilian Casualties from the El Salvadoran Civil War}
\label{sec:case-studies}

In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of individuals recorded in each dataset. Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors. This case study also considers the ramifications of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.

\subsection{El Salvadoran Civil War}
\label{el_salvador}

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualties, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casualty lists. This is an important step in estimating the total number of deaths from a conflict using multiple systems estimation \citep{lum2013applications} Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
two different sources, namely,  El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized denunciations that
had been published throughout the conflict, and additional investigation was required before the ERTL recorded them as human rights abuses, giving us confidence in its accuracy \cite{howland2008rescate}. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized; since reports were taken soon after the events occured, this dataset is also thought to be fairly reliable. \cite{ball2000salvadoran}.

There are several challenges with working with such data. Firstly, both
datasets have been automatically digitized, which inherently leads to
some degree of typographical error. Secondly, the CDHES records are all
second hand accounts reported by individuals, which can result in
additional errors. Lastly, the only fields recorded are given name, last
name, date of death, and place of death; it is relatively common for a
parent and child to share the same given name, resulting in
indistinguishable records for two different individuals. This last point
nearly breaks the earlier mentioned assumption that there are no
duplicates within files, and reveals a key difference between
\texttt{BRL} and our proposed method.

In this paper, we utilize records that have non-missing entries for given and last name, which results 
in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
in ERTL. We standardized names to account for common misspellings
in the Spanish language. The string distance used is a modified Levenstein
distance to account for the fact that second names are often omitted.
Place of death is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
Thus, we conduct record linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.
\begin{table}
\begin{tabular}[h!]{ll|llll}
	\hline
	\multicolumn{2}{c|}{ } & \multicolumn{4}{c}{Level of Disagreement} \\
	\cline{3-6}
	Fields & Similarity & 1 & 2 & 3 & 4\\
	\hline
	First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, .5] & (5, 1]\\
	\hline
	Year of Death & Absolute Difference & 0 & 1 & 2 & 3+\\
	\hline
	Month of Death & Absolute Difference & 0 & 1 & 2-3 & 4+\\
	\hline
	Day of Death & Absolute Difference & 0 & 1-2 & 3-7 & 8+\\
	\hline
	Municipality & Binary & Agree & Disagree &  & \\
	\hline
\end{tabular}
\caption{Construction of comparison vectors for El Salvador data resembling original implementation from \citep{sadinle2018bayesian}. This set up leads to 2048 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_1}
\end{table}

\vspace{1em}

\begin{table}
	\centering
	\begin{tabular}[h!]{ll|lll}
		\hline
		\multicolumn{2}{c|}{ } & \multicolumn{3}{c}{Level of Disagreement} \\
		\cline{3-5}
		Fields & Similarity & 1 & 2 & 3\\
		\hline
		First and Last Name & Modified Levenstein & 0 & (0, .25] & (.25, 1]\\
		\hline
		Year of Death & Binary & Agree & Disagree & \\
		\hline
		Month of Death & Binary & Agree & Disagree & \\
		\hline
		Day of Death & Absolute Difference & 0 & 1 & 2+\\
		\hline
		Municipality & Binary & Agree & Disagree & \\
		\hline
	\end{tabular}
\caption{Construction of comparison vectors for El Salvador for increased speed under \texttt{fabl}. This set up leads to 216 possible agreement patterns in total}\label{Tab:el_salvador_cutoffs_2}
\end{table}

To mirror the original implementation of \cite{sadinle_bayesian_2017}, we construct the comparison
vectors using four levels of agreement for each field, according to the
thresholds provided in Table~\ref{Tab:el_salvador_cutoffs_1}. Gibbs sampling for these comparison vectors took 422 seconds for our proposed
method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
agreement levels creates unnecessary distinctions in the data and makes
the comparison vectors less interpretable.

Therefore we re-ran our analysis with fewer agreement levels for each field according to Table~\ref{Tab:el_salvador_cutoffs_2}, and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method, \texttt{fabl} became much faster,
finishing in 61 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computational
gains can be made by simplifying the construction of the comparison
vectors. The estimates of the \(\mathbf{m}\) parameters under each method are very similar, as shown in
Figure~\ref{fig:m-and-u}.

Through \texttt{fabl}, we arrive at a Bayes estimate of 179 individuals recorded in both datasets. We calcuate posterior samples of size of the overlap across files by summing the number of matches found in each iteration of the Gibbs sampler, and subtracting the number of matches that violate one-to-one matching. The posterior 95\% credible interval for the overlap across files is (206, 238), indicating that the Bayes estimate is quite conservative. This is because there a large number of records in ERTL for which there are multiple plausible matches in CDHES; \texttt{fabl} recognizes that a match exists among the several options, but is unable to definitely declare a specific pair a match in the Bayes estimate. We also compute a partial estimate of the linkage structure, using $\theta_{10} = \theta_{01} = 1$, $\theta_{11} = 2$, and $\theta_R = 0.1$ as in the simulation study. Here, the Bayes estimate provides 136 matches of which the model is quite confident, and 175 records to verify manually; this means that after clerical review, the number of individuals replicated across datasets would fall in the interval (136, 311), encapsulating the posterior credible interval. More or fewer records could be indentified for clerical review by decreasing or increasing $\theta_R$.

We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
		\caption{Posterior distribution and Bayes estimate of overlap across the two files. We note they are quite similar under both methods.}
		\label{fig:overlap-plot}
	\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
\caption{Posterior estimates of m parameters with 95\% credible intervals. We note they are quite similar across the two methods. }\label{fig:m-and-u}
\label{fig:m-and-u}
\end{center}
\end{figure}

\paragraph{Violations of One-to-One Matching}

As noted previously, the application exhibits many situations in which bipartite matching is difficult. To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation. These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in \(\bm{X}_1\) has
multiple plausible matches in \(\bm{X}_2\).

%Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
%duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
%see that \texttt{fabl} consistently overmatches within each Gibbs
%iteration when compared to \texttt{BRL}, which is to be expected because
%\texttt{BRL} explicitly prevents matches that violate one-to-one
%matching throughout the entire sampler. Most of such matchings are due
%to the randomness in the sampling procedure, and they occur sporadically
%throughout the sampler in such a way that does not measurably influence
%the eventual Bayes estimate.

Figure~\ref{Tab:rosa-maria} illustrates an example. Note that these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

\begin{table}
	\centering
	\begin{tabular}[h!]{llllllll}
		\hline
		Record Number & dataset & lastname & firstname & day & month & year & dept \\
		\hline
		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
		\hline
		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
		\hline
		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & NA\\
		\hline
	\end{tabular}
	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.} \label{Tab:rosa_maria}
\end{table}

Figure~\ref{fig:mixing-plot} shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(\bm{X}_2\) match to the
same record in \(\bm{X}_1\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration and stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. In addition,
\texttt{fabl} allows the modeler to inspect records with multiple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate. In short, we have pointed out why one may not wish to imbed this constraint in a model, and proposed a simple alternative that performs about the same as \texttt{BRL}, while scaling to much larger data sets.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
\caption{Gibbs sampling in situation with multiple plausible matches. Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability. Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
\end{center}
\end{figure}

%\begin{figure}[t]
%	
%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
%		
%	}
%	
%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
%\end{figure}

%\subsection{National Long Term Care Survey}
%\label{nltcs} 
%
%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients. It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989. The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office. State and regional office have been anonymized, presented as numbers instead of their actual values. We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context. The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns. We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.  
%
%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible. Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold. Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets. We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing. We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler. 
%
%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$. However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds. Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching. 
%
%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking. With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks. This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler. However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error. In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office. We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives. Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures. However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons. Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block. Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.

\section{Discussion}
\label{discussion}

We have presented a method for Bayesian record linkage that is feasible
for large datasets. In particular, our hashing procedure and model
assumptions allow for a linkage procedure whose computational complexity
does not scale with the size of the larger dataset, making this method
particularly powerful in linking records when one datafile is
substantially smaller than the other.

In our case study, we included an exploration of how to conduct record
linkage when modeling assumptions are not met in practice. We explored
``one-to-many'' scenarios in which one record in \(A\) has multiple
plausible matches in \(B\), and showed how both \texttt{fabl} and
\texttt{BRL} demonstrated undesirable qualities. Other issues arise
under ``many-to-one'' scenarios, where one record in \(B\) has multiple
plausible matches in \(A\), and ``many-to-many'' scenarios in which
there is duplication both across and within datasets. Tuning
\texttt{fabl} for use in these scenarios is one potential avenue for
future work.

\clearpage

\section{Appendix}
\label{sec:appendix}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_trace} 
\caption{Traceplot for number of matches found across datasets in El Salvador case study}\label{fig:overlap_trace}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_trace} 
\caption{Traceplot for m parameter in El Salvador case study}\label{fig:m_trace}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/u_trace} 
\caption{Traceplot for u parameter in El Salvador case study}\label{fig:u_trace}
\end{center}
\end{figure}

\clearpage

\bigskip

\bibliographystyle{jasa}
{
\footnotesize
\bibliography{biblio}
}

\end{document}