### 2022-07-14 methods_paper/cover-letter/ba-cover-letter.tex

\documentclass[11pt]{letter}
%\usepackage{fullpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[top=1in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
                      % LOOKING THE WAY YOU WANT IT
	bottom=1in, left=1in, right=1in]{geometry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{eso-pic}
%\usepackage{graphicx}
\usepackage{parskip}

\newcommand\BackgroundPic{
\put(0,0){
\parbox[b][\paperheight]{\paperwidth}{
\vfill
\centering
\includegraphics[width=\paperwidth,height=\paperheight,
keepaspectratio]{beka_lh_duke.pdf}%
\vfill
}}}

%\setlength\headheight{25em}
%\setlength\headsep{0em}
%\date{June 9, 2010}
%\date{\raggedright November 29, 2021}

\begin{document}
\date{\raggedright December 21, 2021}
%\AddToShipoutPicture*{\BackgroundPic}
%\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
%\thispagestyle{empty}

\begin{letter}{
%Dear Graduate Admissions Committee
%https://www.franklin.uga.edu/jobs/statistics
%National Science Foundation Graduate Research Fellowship
%Carnegie Mellon University\\Department of Statistics\\
%Baker/Porter Hall\\
%Pittsburgh, PA 15213}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
                % CHANGE THIS NUMBER TO GET THE FIRST PAGE HOW YOU WANT IT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{left}
%This goes left\\
%BxLah\\
%\end{left}

\opening{Dear Editor Guindani:}

I am a PhD student at Duke University in the Department of Statistical Science. I am writing to submit my paper with my coauthors
Jerome P. Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.

Very often information about social entities is scattered across multiple databases.  Combining that information into one database can result in enormous benefits for analysis, resulting in richer and more reliable conclusions. The application areas touched by record linkage include voter registration, official statistics, human rights violations, patent inventions, health applications, among many others. Disciplines that utilize both applications and methods include statistics, biostatistics, computer science, machine learning, political science, and others.

Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or fabl. Specifically, in fabl we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that fabl has markedly increased speed with minimal loss of accuracy.

To help the editorial office, we have listed a number of potential reviewers:

\setlength{\parindent}{0mm}
\hangindent=10mm
\hangafter=0
Mauricio Sadinle, msadinle@uw.edu \newline
Serge Aleshin-Guendel, aleshing@uw.edu \newline
Yves Thibaudeau, Yves.Thibaudeau@census.gov \newline
Megan Price, meganp@hrdag.org \newline
Ted Enamorado, ted@wustl.edu \newline
Nick Monath, nmonath@cs.umass.edu \newline
Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au \newline
Ben Rubinstein, benjamin.rubinstein@unimelb.edu.au \newline
Neil Marchant, nmarchant@unimelb.edu.au \newline
Andy Toole, Andrew.Toole@uspto.gov \newline
Giacomo Zanella, zanella.gcm@gmail.com \newline
Brunero Liseo, brunero.liseo@uniroma1.it \newline
Andrea Tancredi, andrea.tancredi@uniroma1.it \newline
Peter Christen, peter.christen@anu.edu.au \newline
Partha Lahiri, plahiri@umd.edu \newline
Rebecca Nugent, rnugent@stat.cmu.edu \newline
Kayla Frisoli, kaylafrisoli@gmail.com \newline

We appreciate consideration of our paper, and please do not hesitate to reach out if you have any questions.

 \closing{Sincerely,}
 Brian Kundinger

\end{letter}

\end{document}