### 2023-06-07 Prelim/bcm_proposal.tex
corrected SEI

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }

\begin{document}
	\title{Bayesian Causal Matching with Fellegi Sunter}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger}
	\institute{Duke University}
	\date{\today}

%	\AtBeginSection[]{
%		\begin{frame}
%			\vfill
%			\centering
%			\begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
%				\usebeamerfont{title}\insertsectionhead\par%
%			\end{beamercolorbox}
%			\vfill
%		\end{frame}
%	}

\AtBeginSection[]
{
	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents[currentsection]
	\end{frame}
}

	\begin{frame}
		\titlepage
	\end{frame}

\section{Matching in Causal Inference}
\begin{frame}{Overview of Approaches}
	\begin{itemize}
		\item Exact Matching
		\item Mahalonobis Distance
		\item Propensity Score
		\item Almost Exact Matching
	\end{itemize}
\end{frame}

\begin{frame}{Mahalonobis Distance}
	\begin{itemize}
	\item One method of matching has been to calculate distance in the covariate space through
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1} |X_1 - X_2|}
		\end{align}
	\item An algorithm then matches each unit in control to closest unit in treated
	\end{itemize}
\end{frame}

\begin{frame}{Propensity Score}
	\begin{itemize}
	\item Another method is to estimate propensity scores $p(Z_n = 1 |X_n)$ for each unit in the data, and match units based on this score.
	\end{itemize}
\end{frame}

\begin{frame}{Pitfalls of Both Approaches}
	\begin{itemize}
		\item The pros and cons of each method are summarized very well on this \href{https://stats.stackexchange.com/questions/511294/what-are-the-pros-and-cons-of-using-mahalanobis-distance-instead-of-propensity-s}{Stack Exchange} page

		\item Essentially, propensity scores can end up matching units with that have similar probability of treatment, but have very different covariate values
		\item Conversely, Mahalonobis can end up matching units that are similar on features that are irrelevant to the outcome of interest
	\end{itemize}
\end{frame}

\begin{frame}{Blending the Two Approaches}
	\begin{itemize}
		\item There is also a good summary on page 5 of \href{https://www.jstatsoft.org/article/view/v042i07}{Sekhot (2011)} recommends including the propensity score into the Malahalonobis distance. This was the genesis for the idea to come
	\end{itemize}
\end{frame}

\section{Proposed Idea}

\begin{frame}{My Idea}
	\begin{itemize}
		\item The machinery of Fellegi Sunter record linkage is very good at identifying duplicate records. It should also be good at identifying separate individuals that are similar.
		\item FS does matching purely on the reliability and uniqueness ($m$ and $u$ parameters) of fields of information. It does not take into account any relationship with an outcome variable. Conceptually, this is very similar to the Mahalonobis distance, but allows for uncertainty quantification.
		\item If we include propensity scores into FS, the model will place high weight on units with ``similar" scores. Then, the FS machinery for the other fields should find the units that are most similar. 
	\end{itemize}
\end{frame}

\begin{frame}{A Similar Idea}
	\begin{itemize}
		\item\href{https://arxiv.org/pdf/2105.02362.pdf}{Alvarez and Levin (2021)} (Section 2, pages 5 - 7) use a Bayesian model to create draws of the propensity scores. For set of scores, they use a deterministic algorithm to match records. 
		\item I imagine that each particular draw of matched pairs may have the issues presented earlier. Propensities are matched, but units may actually look quite different.
		\item Information does not seem to propagate correctly through the Gibbs Sampler
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Data is $\{y_n, z_n, x_n\}_{n = 1}^N$ where $y_n$ is the outcome, $z_n$ is the treatment indicator, and $x_n$ is a vector of $F$ fields of covariates. 
		\item Split data into a datasets $T$ with records $i \in \{1, \ldots, n_T\}$ for treated units and $C$ with records $j \in \{1, \ldots, n_C\}$. 
		\item At each step of the Gibbs Sampler, calculate propensity score $PS(i) = P(Z_i = 1 | X)$ for each data point (logistic or probit regression)
	\end{itemize}
\end{frame}

\begin{frame}{My Idea}
	\begin{itemize}
		\item Define comparison field 
		\begin{align}
			\Gamma_{0} = \begin{cases}
				1 & $PS(i)$ and $PS(j)$ \text{ are sufficently close} \\
				0 & \text{otherwise} \\
			\end{cases}
		\end{align} 
		\item ''Sufficiently close" could mean all units with scores within a certain caliper, or it could mean the $K$ nearest scores.
		\item Conduct Fellegi Sunter Matching on new full set of comparison vectors $\Gamma^{*} = (\Gamma_0, \gamma)$.
		Note $\gamma$ is fixed data, but the $\Gamma$ are random. 
	\end{itemize}
\end{frame}

\begin{frame}
	\begin{itemize}
		\item Each iteration of the Gibbs sampler will provide a matched dataset, from which you can calculate an MCMC sample of the treatment effect (TE) or other quantities of interest
		\item We obtain measures of match quality as usual, through the posterior samples of the linkage structure
		\item I believe this to be first causal matching method to produce matchings probabilistically
		\item The proposed method is fully model based, with no deterministic algorithms. It propogates uncertainty between the matching procedure and the treatment effect estimation. 
	\end{itemize}
\end{frame}

\begin{frame}{Model (Basic)}
	\begin{align}
		Z_n &= x_n' \beta + \epsilon \\
		\beta &\sim N(0, \tau) \\
		\epsilon &\sim N(0, \sigma^2) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l) \\
		m, u \sim \text{Dirichlet}
		\Delta \sim \text{Desired Record Linkage Prior}
	\end{align}
\end{frame}

\begin{frame}{Model (With Variable Selection)}
	\begin{itemize}
		\item In causal matching, it is especially important not to match on irrelevant features. Therefore, we can use a spike and slab prior on $\beta$, and only use fields where $\beta_f \neq 0$ in the linkage.
	\end{itemize}
\begin{align}
	\beta_j &\sim \pi_0 \delta_0 + \(1 - \pi\)N(0, \tau_j) \\
	\mathcal{L}(\Delta, m, u \mid \Gamma^{*}, \beta) &= \prod_{i=1}^{n_T}  \prod_{j=1}^{n_C}\prod_{f=0}^{F}\left[\prod_{l=1}^{L_f}\left[  m_{fl}^{\Delta_{ij} = 1}u_{fl}^{\Delta_{ij} = 0}\right]^{I(\gamma_{ij}^f = l)\right]^{\beta_j \neq 0} \\
\end{align}
\end{frame}

\begin{frame}{Full Conditional for $\beta$}
	\begin{itemize}
		\item The full condition for $\beta_j$ includes a measure of the quality of matches along field $f$
	\end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}
	\begin{itemize}
		\item Mathematically, any record linkage prior could work. Sadinle, fabl, or original FS (one-to-one, one-to-many, many-to-many) may all be useful , depending on the analysis
		\item If this works, would be a very strong final thesis chapter
	\end{itemize}
\end{frame}

\begin{frame}{Caveats}
	\begin{itemize}
		\item Using a comparison vector made from the propensity score may technically violate the conditional independence assumption of FS
		\item I can't imagine this idea totally failing, but I'm sure that it won't be cutting edge. Sekhot further devised a method that estimates
		\begin{align}
			d(X_1, X_2) = \sqrt{|X_1 - X_2| S^{-1/2} W S^{-1/2} |X_1 - X_2|}
		\end{align}
		where the entries in $W$ are fine tuned through optimization methods that encourage covariate balance across datasets, based on t-tests and all sorts of other measures. This is almost certainly outperform what I have presented, because it is able to use more information.
		\item Propensity and Mahalonobis matching both have theoretic garuntees. I don't know right now if this method would have them
		\item I know very little about causal inference! This would be a large undertaking
	\end{itemize}
\end{frame}

\begin{frame}{Questions}
	\begin{itemize}
		\item What do you all think of this idea?
		\item Jerry, how well versed are you in matching for causal inference? To what extent to should I seek guidance/collaboration from Fan Li or Alex Volfovsky?
	\end{itemize}
\end{frame}

\begin{frame}{Possible Theoretical Question}
	\begin{itemize}
		\item I'm really struck by the conceptual similarity between Mahalanobis distance and the FS weights. If you make some assumptions (categorical data, one-to-one matching), it might be possible to prove a correspondence between the two. 
	\end{itemize}
\end{frame}

\end{document}