### 2023-07-19 SEI_sensitivity/Sadinle_accuracy_cluster.Rmd
pi in gibbs sampler

---
title: "Sadinle_accuracy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sadinle_accuracy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(devtools)
load_all()
library(parlrdev)
library(tidyverse)
rm(list = c("hash_field"))
```

```{r}
taskID <- as.integer(Sys.getenv("SLURM_ARRAY_TASK_ID"))
i = taskID
set.seed(41)
files <- list.files(path = "../data/SimulationDataFiles", full.names = T)

m_prior = 1
u_prior = 1
alpha = 1
beta = 1
S = 1000
burn = 100
show_progress = T
fast = F
R = NULL
all_patterns = TRUE

overlap_vals <- c(50, 250, 450)

fabl_acc_samps <- matrix(NA, nrow = 5, ncol = 3)
fabl_acc_samps_resolved <- matrix(NA, nrow = 5, ncol = 3)
sad_acc_samps <- matrix(NA, nrow = 5, ncol = 3)

fabl_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)
fabl_acc_samps_resolved2 <- matrix(NA, nrow = 6, ncol = 3)
sad_acc_samps2 <- matrix(NA, nrow = 6, ncol = 3)

for(j in seq_along(overlap_vals)){

  overlap <- overlap_vals[j]

#for(i in seq_along(files)){

records <- read_csv(files[i], col_types = cols())
records$file <- rep(2:1, length.out = dim(records)[1])

records <- records %>% 
  janitor::clean_names() %>% 
  mutate(rec_id = as.numeric(str_extract(rec_id, "\\d{3}")) + 1)

n1 <- 500
n2 <- 500
#overlap <- n2/2

Ztrue <- n1 + 1:n2
Ztrue[1:overlap] <- 1:overlap

file1 <- records %>% 
  filter(file ==1,
         rec_id <= n1) %>% 
  select(-rec_id) %>% 
  as.matrix(.) %>% 
  data.frame(.) %>% 
  mutate(occup = as.numeric(occup))

file2 <- records %>% 
  filter(file == 2,
         rec_id %in% c(1:overlap, (n1 +1):(1000 - overlap))) %>% 
  select(-rec_id) %>% 
  as.matrix() %>% 
  data.frame(.) %>% 
  mutate(occup = as.numeric(occup))

cd <- suppressMessages(BRL::compareRecords(file1, file2, c(2, 3, 5, 6), 
                          types = c("lv", "lv", "bi", "bi")))
cd[[1]] <- apply(cd[[1]], 2, as.numeric)

  #parlr method
  ptm <- proc.time()
  Zchain <- BKSimple_hash2(cd, R = R)
  Zchain <- fabl_gibbs(cd, R = R)
  elapsed <- proc.time() - ptm
  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, Inf)
  eval <- GetEvaluations(Zhat[[1]], Ztrue, n1)
  eval

  # Resolution Step
  Zhat_resolved <- ResolveConflicts(Zhat)
  eval_resolved <- GetEvaluations(Zhat_resolved, Ztrue, n1)
  fabl_acc_samps_resolved[, j] <- c(eval_resolved, elapsed[3], overlap)
  fabl_acc_samps[, j] <- c(eval, elapsed[3], overlap)

  #Partial Estimate
  Zhat <- LinkRecordsBK(Zchain[[1]], n1, 1, 1, 2, .1)
  Zhat <- ResolveConflicts(Zhat)
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  RR <- sum(Zhat == -1)/n2
  eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
  fabl_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)

  #Sadinle 2017 Method
  ptm <- proc.time()
  Zchain <- BRL::bipartiteGibbs(cd)[[1]]
  Zchain <- Zchain[,101:1000]
  elapsed <- proc.time() - ptm
  Zhat <- BRL::linkRecords(Zchain, n1, 1, 1, 2, Inf)
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  sad_acc_samps[, j] <- c(eval, elapsed[3], overlap)

  Zhat <- BRL::linkRecords(Zchain, n1, 1, 1, 2, .1)
  RR <- sum(Zhat == -1)/n2
  eval <- GetEvaluations(Zhat, Ztrue, n1)
  eval[1] <- sum(Zhat == Ztrue & Ztrue > n1) / (sum(Zhat > n1))
  sad_acc_samps2[, j] <- c(eval, RR, elapsed[3], overlap)

#  print(i)
#}
}

saveRDS(Zs, file = paste0("out/fabl_acc/fabl_acc_", 
                          str_pad(fabl_acc_samps, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/fabl_acc_res/fabl_acc_res_", 
                          str_pad(fabl_acc_samps_resolved, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/sad_acc/sad_acc_", 
                          str_pad(sad_acc_samps, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/fabl_acc2/fabl_acc2_", 
                          str_pad(fabl_acc_samps2, 3, pad = "0")))
saveRDS(Zs, file = paste0("out/sad_acc2/sad_acc2_", 
                          str_pad(sad_acc_samps2, 3, pad = "0")))
# list(BK = parlr_acc_samps,
#      Sadinle = sad_acc_samps)

```